\textbf{ Spectrophotometric parallaxes with linear models:\\ Accurate distances for luminous red-giant stars }

### 2019-06-02 paper/response_to_referee.txt

To-do items:
------------
- find and kill all HOGG
- find and kill all XXX

@@

    The paper presents a method to estimate spectrophotometric parallaxes
    based only on input data, without adoption of physically motivated
    models and priors. The paper is written in an way that is strongly
    focused on the method itself, but leaves several complementary aspects
    that are certainly important to assess the reliability of the inferred
    parallaxes and, most importantly, distances. I recommend its
    publication only after such major issues are addressed.

    Major issues

    1 - The data-driven method described in the paper details out how
    spectrophotometric parallaxes are inferred. But the paper does not
    give the reader a clue about how spectrophotometric distances are
    estimated. The authors make it clear in Section 2 that they never
    invert *observed* parallaxes to estimate distances in their
    method. But are the spectrophotometric distances they infer (shown in
    Figures 5 and 6) just the inverse of the *inferred* parallaxes? If so,
    this should be made explicit in the text, along with some discussion
    about the problems (or lack of) that may arise from this
    inversion. Even though the spectrophotometric parallaxes themselves
    are always positive, a reader would like to be convinced that the
    distances shown in Figures 5 and 6 do not suffer from biasses caused
    by the asymmetric nature of the distances distribution that results
    from a symmetric error distribution in parallaxes (as stated by the
    likelihood assumption of the method). How do these biasses simply go
    away when the method's w(sp) +/- sigma(sp) are used? In particular, a
    negative parallax for many stars may be consistent with their w(sp)
    +/- sigma(sp).

These issues are subtle! We are assuming that the GAIA-MEASURED
PARALLAXES have noise that is symmetric in parallax space. In our
data-driven model, we use a linear combination of features to predict
or estimate a LOG PARALLAX, so if anything we are claiming that our
estimates (not the data, but our estimates based on the data) are
symmetric in log parallax, not parallax. But we aren't even really
claiming that. The issues are subtle. However, this does deserve some
discussion and we have added that to the captions of Figure XXX and
XXX (formerly Figure 5 and 6), and in the last paragraph of the
discussion, for extra-special emphasis.  The short answer is that we
just exponentiated the negative of our log-parallax estimates to make
those figures.

@@

There is an additional question of whether we are somehow biasing
things when we presume that the parallax must be positive? Our
position is that we are biasing things by presuming that the LOG
PARALLAX can be predicted as a linear combination of features. Of
course any other assumption would also introduce biases! There is no
way to fit that would be unbiased here, because by construction, our
model is not physically accurate. We added words about this into the
linearity assumption in the assumption list in Section 2.

    2 - The paper shows limited validation tests and comparisons to other
    work. There are several papers that provide spectrophotometric
    distances based on physical models for APOGEE stars (Queiroz et al
    2018, MNRAS, 476, 2556; Schultheis et al 2014, AJ, 148, 24 and
    others). The authors should compare their distances to those authors
    and discuss the results of this comparison. Moreover, asteroseismic
    distance estimates are also available for a fraction of the APOGEE
    sample of giants in common with the CoRoT and Kepler missions. This is
    an additional and important validation test.

This is a great point. We obtained these two data sets (and the
Queiroz sample required some emails!) and we made comparisons. We
added a short paragraph about the results of these comparisons in the
second-last paragraph of the Results Section. We did not feel that the
comparisons merited supporting Figures, but we do have such figures if
the referee disagrees.

During the long period in which we haven't responded to this report,
there was also a paper out by Leung & Bovy. They compare their results
to ours, and we also mention that in the newly added paragraph.

### 2019-06-02 paper/sp.tex
another point down

@@

>>>>

It is also interesting to compare these estimates to other, previous
spectroscopic distance efforts.
We performed comparisons with two other efforts (\citealt{sch, que}).
In both cases, the agreement is better at smaller distances (larger
parallaxes) and worse at larger distances, with the distances
(parallaxes) presented here being smaller (larger) on average than the other
investigation results for the very most distant (and hence most difficult) stars.
In detail, the comparison with the larger sample (\citealt{sch})
are better:
The scatter is at the 15-percent level for stars at distances
$<8$\,kpc), and there is no visible bias between the results at these
distances.
The distances presented here also compare well with those from a recent
machine-learning approach (\citealt{leung}), which appeared while this
paper was under review.

@@

\bibitem[Queiroz \etal(2018)]{que} Queiroz, A.~B.~A., Anders, F., Santiago, B.~X., \etal\ 2018, \mnras, 476, 2556.
\bibitem[Schultheis \etal(2014)]{sch} Schultheis, M., Zasowski, G., Allende Prieto, C., \etal\ 2014, \aj, 148, 24
